Long name of the paper

### 2018-08-30 anzjs2018/sections/03_models.tex
words

The first of these is incorporated into $\bM_1$,
which allows vehicle speeds to vary over time.
This model uses the same state as $\bM_0$,
but a different transition function $f_1$.
In this case, the state equations depend on a sequence of system noise,
$\bw_k = \{\bw_k^1, \cdots, \bw_k^{\Delta_k}\}$
\begin{align*}
x_{k|k-1} &= x_{k-1|k-1} + \Delta_k \dot x_{k-1|k-1} + \sum_{i=1}^{\Delta_k} w_k^i \\
\dot x_{k|k-1} &= \dot x_{k-1|k-1} + \sum_{i=1}^{\Delta_k} w_k^i \\
w_k^i &\sim \mathcal{N}_T(0, \sigma_x^2)
\end{align*}
Again, system noise is assumed to have a Gaussian distribution, 
truncated to ensure that the vehicle's speed is always between 0 and 30~m/s.
This is achieved by iteratively estimating each second the the vehicle's state,
each time ensuring the speed remains in the specified range.

>>>>

The first of these is incorporated into $\bM_1$,
which allows vehicle speeds to vary over time.
This model uses the same state as $\bM_0$,
but a different transition function $f_1$.
In this case, the state equations depend on a sequence of system noise,
$\bw_k = \{\bw_k^1, \cdots, \bw_k^{\Delta_k}\}$
\begin{align}
\label{eq:transition_f1}
x_{k|k-1} &= x_{k-1|k-1} + \Delta_k \dot x_{k-1|k-1} + \sum_{i=1}^{\Delta_k} w_k^i \nonumber \\
\dot x_{k|k-1} &= \dot x_{k-1|k-1} + \sum_{i=1}^{\Delta_k} w_k^i \\
w_k^i &\sim \mathcal{N}_T(0, \sigma_x^2) \nonumber
\end{align}
Again, system noise is assumed to have a Gaussian distribution, 
truncated to ensure that the vehicle's speed is always between 0 and 30~m/s.
This is achieved by iteratively estimating each second the the vehicle's state,
each time ensuring the speed remains in the specified range.

The next model, $\bM_2$, allows for dwell times at bus stops.
Since buses can either stop and wait, or not stop,
multimodality is introduced to the system,
which requires a modeling framework, such as the particle filter used here,
to handle it.
For each bus stop $j = 1, \ldots, M$ along a route,
its distance into the trip, $S_j^d$, 
is used to determine when a vehicle passes a stop.
When this occurs, the dwell time $\delta_j \geq 0$ at stop $j$ is modelled
using the stopping probability, $\pi_j$,
a common dwell time parameter $\gamma$ which models the
deceleration, opening and closing of doors, and acceleration at bus stops,
and a stop dwell time $\tau_j$ for passengers alighting and disembarking.
Therefore the stop dwell time is
\begin{equation}
\label{eq:dwell_time}
\delta_j = p_j(\gamma + \tilde\delta_j),\quad
p_j \sim \mathrm{Bern}(\pi_j),\quad
\tilde\delta_j \sim \mathcal{E}(\tau_j)
\end{equation}

To define the transition function for this model, $f_2$,
we need to account for multiple dwell times.
Let $s_n$ be the index of the next stop,
and sample dwell times for each upcoming stop $j = s_n, \ldots, M$,
noting that these can be zero.
Assuming the vehicle passes stops $s_n$ up to $J-1$,
does it reach stop $J$ in the time remaining?
\begin{equation}
\tilde X_k^J = \min\left\{
S_J^d, f_1(\hat X_{k|k-1}, \Delta_k - \sum_{j=s_n}^{J-1}\delta_j)\right\}
\end{equation}
Taking the maximum of the sequence gives the predicted distance
traveled by the vehicle, so
\begin{equation}
\hat X_{k|k-1} = f_2 (\hat X_{k-1|k-1}, \Delta_k) =
\max_{J = s_n, \ldots, M}\left\{
    \tilde X_k^J
\right\}
\end{equation}

Care needs to be made when generating the sequence,
so in practice the state will be calculated recursively to ensure
the system noise is consistent.
\begin{equation*}
\hat{\tilde X}_k^J =
\begin{cases}
    f_1(\hat{\tilde X}_k^{J+1}, \delta_J) & J < M \\
    \hat X_{k-1|k-1} & J = M
\end{cases}
\end{equation*}
